s.function

### 2023-12-12 src/la_RF.Rmd
CAST
refined model LA

@@

```{r}
# get LA for the UK
path.geo <- paste0(path, "/data/raw/Local_Authority_Districts_(December_2021)_UK_BUC.geojson")
la <- readOGR(path.geo)
# source: https://geoportal.statistics.gov.uk/

@@

```{r}
# LA centroids
la.c <- gCentroid(la, byid=TRUE)@coords %>% 
  as.data.frame() %>%
  bind_cols(la@data) %>% 
  dplyr::select(LAD21CD, x, y) %>% 
  remove_rownames()

@@

```{r}
dist <- distm(cbind(la.c$x, la.c$y), cbind(cities$long, cities$lat), fun=distHaversine) 
dist <- round((dist/1000),2) %>% 
  as_tibble()

@@

# calculate area
la$SHAPE_Area <- area(la) #/ 1000000

# Join with distance and area
df <- df %>% left_join(dist, by = c("ladcd" = "ladcd")) %>% 
  left_join(la@data, by = c("ladcd" = "LAD21CD")) %>% 
  rename(area = SHAPE_Area) %>% 
  relocate(area, .after = n) %>% 
  mutate(area = area / 1000000,
         density = n / area) %>% 
  relocate(density, .after = area)
sapply(df, function(x) sum(is.na(x)))
```

@@

```{r}
dist.retail <- distm(cbind(la.c$x, la.c$y), cbind(retail.major.cetres$X, retail.major.cetres$Y), fun=distHaversine) 
dist.retail <- round((dist.retail/1000),2) %>% 
  as_tibble()

@@

```{r}
# London = Islington 
sf_use_s2(FALSE)
path.geo <- paste0(path, "/data/raw/Local_Authority_Districts_(December_2021)_UK_BUC.geojson")
la.sf <- st_read(path.geo)

help1 <- st_filter( la.sf, cities.sf) %>% as_tibble %>% dplyr::select(LAD21CD) %>% 
  left_join(df, by = c("LAD21CD"="ladcd")) %>% dplyr::select(LAD21CD, year, n)

help1 <- st_filter( la.sf, cities.sf) %>% as_tibble %>% dplyr::select(LAD21NM,LAD21CD) %>% 
  mutate(cities = c("Leicester",
                    "Bristol",
                    "Manchester",
                    "Liverpool",
                    "Sheffield",
                    "Birmingham",
                    "Leeds",
                    "London",
                    "Edinburgh",
                    "Glasgow"))

@@

sf_use_s2(FALSE)
la.c <- st_centroid(la.sf)
kn <- knearneigh(la.c, k = 5)
knn <- knn2nb(kn, row.names = NULL, sym = FALSE)
knn.l <- nb2listw(knn)

df <- df %>% drop_na() %>% group_by(year) %>% 
  mutate(n.slag = lag.listw(knn.l, n)) %>% 
  group_by(ladcd) %>% 
  mutate(n.l.slag = dplyr::lag(n.slag, n=1, order_by=year),
         n.lag = dplyr::lag(n, n=1, order_by = year)) %>% 
  left_join(help2, by = c("ladcd" = "ladcd"))

@@

# For growth:
#df <- df %>% mutate(growth = log(n/n.lag))

@@

df <- df %>% filter(year > 1996)

#df <- df %>% dplyr::select(-n.lag)
```

@@

```{r}
set.seed(123)
indices <- CreateSpacetimeFolds(df, 
                                spacevar = "ladcd",
                                timevar = "year", 
                                k = 10)

@@

#(8 is the number of tuning parameter, mtry for rf, here equal to ncol(iris)-2)
for(i in 1:10) seeds[[i]]<- sample.int(n=1000, 11) # 9 after adding South

@@

start.time <- Sys.time()
model.all <- train(n ~ area + dist + dist.retail + London + south + 
                     n.slag + n.lag + n.l.slag + n.nearest.city + year,
                   data = df, 
                   trControl = tc,
                   method = "ranger",
                   na.action = na.omit,
                   #preProc = c("center", "scale"),
                   importance = "impurity")

@@

### train the model in a loop per region

@@

@@

  set.seed(71)
  indices <- CreateSpacetimeFolds(df %>% filter(RGN21CD==i), 
                                #spacevar = "ladcd",
                                timevar = "year", 
                                k = 10)

  model.all <- train(n ~ area + dist + dist.retail + London + south + 
                     n.slag + n.lag + n.l.slag + n.nearest.city + year,
                   data = df %>% filter(RGN21CD==i), 
                   trControl = tc,
                   method = "ranger",
                   na.action = na.omit,
                   #preProc = c("center", "scale"),
                   importance = "impurity")
  assign(paste0("model.all",as.character(i)), model.all)
}

stopCluster(cl)
end.time <- Sys.time()
time.taken <- round(end.time - start.time,2)
time.taken
```

@@

### Test on all other regions

@@

```{r}
# train on one region, test on all the rest

>>>>

## spatial data NEW

```{r}
# get LA for the UK
path.geo <- paste0(path, "/data/raw/Local_Authority_Districts_(December_2021)_UK_BUC.geojson")
la <- st_read(path.geo)
# source: https://geoportal.statistics.gov.uk/

# spatial transformations
la <- st_transform(la, 4326)

@@

```{r, eval=FALSE}
# get LA for the UK
path.geo <- paste0(path, "/data/raw/Local_Authority_Districts_(December_2021)_UK_BUC.geojson")
la <- readOGR(path.geo)
# source: https://geoportal.statistics.gov.uk/

@@

```{r}
# LA centroids
# la.c <- gCentroid(la, byid=TRUE)@coords %>% 
#   as.data.frame() %>%
#   bind_cols(la@data) %>% 
#   dplyr::select(LAD21CD, x, y) %>% 
#   remove_rownames()

sf_use_s2(FALSE)
la.c <- st_centroid(la) %>% 
  st_coordinates() %>%
  as.data.frame() %>% 
  bind_cols(la %>% dplyr::select(LAD21CD))

@@

```{r}
dist <- distm(cbind(la.c$X, la.c$Y), cbind(cities$long, cities$lat), fun=distHaversine) 
dist <- round((dist/1000),2) %>% 
  as_tibble()

@@

# calculate area
#la$SHAPE_Area <- area(la) #/ 1000000
la$area <- st_area(la) #/ 1000000

# Join with distance and area
df <- df %>% left_join(dist, by = c("ladcd" = "ladcd")) %>% 
  left_join(la, by = c("ladcd" = "LAD21CD")) %>% 
  relocate(area, .after = n) %>% 
  mutate(area = area / 1000000,     # UNITS APPEAR WRONGLY IN M AND NOT IN KM
         density = n / area) %>% 
  relocate(density, .after = area)
sapply(df, function(x) sum(is.na(x)))
```

@@

```{r}
dist.retail <- distm(cbind(la.c$X, la.c$Y), cbind(retail.major.cetres$X, retail.major.cetres$Y), fun=distHaversine) 
dist.retail <- round((dist.retail/1000),2) %>% 
  as_tibble()

@@

```{r}
# London = Islington 
# sf_use_s2(FALSE)
# path.geo <- paste0(path, "/data/raw/Local_Authority_Districts_(December_2021)_UK_BUC.geojson")
# la.sf <- st_read(path.geo)

#help1 <- st_filter( la.sf, cities.sf) %>% as_tibble %>% dplyr::select(LAD21CD) %>% 
#  left_join(df, by = c("LAD21CD"="ladcd")) %>% dplyr::select(LAD21CD, year, n)

# nearest city
help1 <- st_filter( la, cities.sf) %>% as_tibble %>% dplyr::select(LAD21NM,LAD21CD) %>% 
  mutate(cities = c("Leicester",
                    "Bristol",
                    "Manchester",
                    "Liverpool",
                    "Sheffield",
                    "Birmingham",
                    "Leeds",
                    "London",
                    "Edinburgh",
                    "Glasgow"))

@@

sf_use_s2(FALSE)
la.c <- st_centroid(la)
kn <- knearneigh(la.c, k = 5)
knn <- knn2nb(kn, row.names = NULL, sym = FALSE)
knn.l <- nb2listw(knn)

df <- df %>% drop_na() %>% # THIS DROPS L999999
  group_by(year) %>% 
  mutate(n.slag = lag.listw(knn.l, n)) %>% 
  group_by(ladcd) %>% 
  mutate(n.l.slag = dplyr::lag(n.slag, n=1, order_by=year),
         n.lag = dplyr::lag(n, n=1, order_by = year)) %>% 
  left_join(help2, by = c("ladcd" = "ladcd"))

@@

# For growth:
df <- df %>% mutate(growth = log(n/n.lag))

@@

df <- df %>% left_join(df %>% ungroup() %>% filter(LAD21NM == "Islington") %>% dplyr::select(year, n) %>% rename(n.London = n)) %>% 
  group_by(ladcd) %>% 
  mutate(n.London.lag = dplyr::lag(n.London)) %>% 
  mutate(n.nearest.city.lag = dplyr::lag(n.nearest.city))

df <- df %>% filter(year > 1996)
sapply(df, function(x) sum(is.na(x)))
```

@@

```{r}
set.seed(123)
indices <- CreateSpacetimeFolds(df %>% filter(year > 1999), 
                                spacevar = "ladcd",
                                timevar = "year", 
                                k = 10)

@@

#(8 is the number of tuning parameter, mtry for rf, here equal to ncol(iris)-2)
for(i in 1:10) seeds[[i]]<- sample.int(n=1000, 6) # , 11) 9 after adding South

@@

start.time <- Sys.time()
model.all <- train(growth ~ #area + dist + dist.retail + London + south + n.slag + n.lag + n.l.slag + London.n.lag + n.nearest.city + year,
                     I(n.London.lag/dist) + I(n.nearest.city.lag/London) + n.l.slag + year,
                   data = df %>% filter(year > 1999), 
                   trControl = tc,
                   method = "ranger",
                   na.action = na.omit,
                   #preProc = c("center", "scale"),
                   importance = "impurity")

@@

### train the model in a loop for all but one regions

@@

# CV
rm(indices)

@@

  set.seed(71)
  indices <- CreateSpacetimeFolds(df %>% filter(RGN21CD!=i), 
                                spacevar = "ladcd",
                                timevar = "year", 
                                k = 10)

  tc <- trainControl(method = "cv", #MOVED HERE< DOUBLE CHECK
                   number = 10,
                   seeds = seeds,
                   allowParallel = T,
                   index = indices$index,
                   savePredictions = 'final')

  model.all <- train(n ~ #area + dist + dist.retail + London + south + n.slag + n.lag + n.l.slag + n.nearest.city + year,
                       I(n.London.lag/dist) + I(n.nearest.city.lag/London) + n.l.slag + year,
                   data = df %>% filter(RGN21CD!=i), 
                   trControl = tc,
                   method = "ranger",
                   na.action = na.omit,
                   #preProc = c("center", "scale"),
                   importance = "impurity")
  assign(paste0("model.all",as.character(i)), model.all) # name = region not in the training data
}

stopCluster(cl)
end.time <- Sys.time() # ca. 2h
time.taken <- round(end.time - start.time,2)
time.taken
```

@@

### Test on one region

@@

```{r}
# train on all but one region, test on that region

df.predictions <- data.frame()
for (i in regions){
    pred <- predict(fit.model.all[names(fit.model.all)==paste0("model.all",i)], df[df$RGN21CD==i,])
    pred <- as.data.frame(pred)
    rownames(pred) <- c()
    assign(paste0("region.predict.model.", ".on.", i), pred)

    d <- paste0("region.predict.model.", ".on.", i)
    d <- cbind(get(d), df %>% filter(RGN21CD==i)) #%>% arrange(year, ladcd)
    d <- d[,c(5, 2, 4, 6, 1)]
    colnames(d)[5] <- "predictions"
    d <- d %>% mutate(tested.on = i)
    df.predictions <- rbind(d, df.predictions)
}

df.predictions <- df.predictions %>% mutate(test.train = paste0(RGN21CD, ".", tested.on))

# split to list by region pair
pred.by.region.all.list <- split(df.predictions, df.predictions$RGN21CD)

path.out <- paste0(path, "/outputs/rf/heatmap_LA.png")
rf.year.all.metrics %>% 
  as.data.frame() %>%
  rownames_to_column(var = "metrics") %>% 
  pivot_longer(!metrics, names_to = "test.on") %>% 
  pivot_wider(names_from = metrics, values_from = value) %>% 
  left_join(lookup.region, by = c("test.on" = "RGN21CD")) %>% 
  arrange(Rsquared)

ggplot(aes(train.region, test.region, fill= Rsquared)) + 
  geom_tile() +
  coord_fixed() +
  labs(y = "", x = "") + #y: test.region, x: train.region
  scale_fill_gradientn(colours = terrain.colors(10), trans = 'reverse') +
  theme(#legend.position = "none",
        axis.text.x = element_text(angle = 45, hjust=1),
        panel.border = element_blank(), 
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_rect(fill = "white"))
ggsave(path.out)

#### OLD

```{r, eval=FALSE}
# train on one region, test on all the rest

### 2023-12-12 src/s_la_per_firm.Rmd
refined model LA

```{r}
a.la <- as_tibble()
all.la <- df %>% filter(ladcd != "NA") %>% distinct(ladcd) %>% 
  #top_n(n=2) %>% 
  simplify2array()
simplify2array(all.la)

@@

a.la <- a.la %>% filter(term=="xmid") %>% 
  bind_cols(as.tibble(all.la)) %>% 
  #mutate(estimate = round(estimate, 0)) %>% 
  #summarise(r = range(r2))
  filter(p.value < 0.01 & 
         r2 > 0.9) %>%    # 376 / 376, all sig, 330 r2 > 0.9
  dplyr::select(ladcd, estimate, r2, std.error) %>%
  mutate(fast = ifelse(estimate < xmid.uk, "fast", "slow")) # it was 2003

>>>>

```{r}
a.la <- tibble()
all.la <- df %>% filter(ladcd != "NA") %>% distinct(ladcd) %>% 
  #top_n(n=2) %>% 
  simplify2array()
simplify2array(all.la)

@@

a.la <- a.la %>% filter(term=="xmid") %>% 
  bind_cols(as_tibble(all.la)) %>% 
  #mutate(estimate = round(estimate, 0)) %>% 
  #summarise(r = range(r2))
  filter(p.value < 0.01 & 
         r2 > 0.9) %>%    # 376 / 376, all sig, 330 r2 > 0.9
  dplyr::select(ladcd, estimate, r2, std.error) %>%
  mutate(fast = ifelse(estimate < xmid.uk, "fast", "slow")) # it was 2003
